Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-10 main.tex
Updates based on yesterday's conversation
and some more questions!

% Notational conventions
\newcommand{\HIGH}{\textsc{h}}
\newcommand{\LOW}{\textsc{l}}
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\PCname(#1)}
\newcommand{\PCname}{\textsc{pc}}
\newcommand{\polname}{P}
\newcommand{\pol}[1]{\polname(#1)}
% \newcommand{\trace}[1]{\overline{#1}}
\newcommand{\contour}{C}
\newcommand{\CONTOURS}{{\mathcal C}}
\newcommand{\component}{K}
\newcommand{\COMPONENTS}{{\mathcal K}}
\newcommand{\stepsto}{\Longrightarrow}
\newcommand{\stepstounder}[1]{\stackrel{\mbox{\tiny{$#1$}}}{\Longrightarrow}}
\newcommand{\stepstounderfull}{\stepstounder{\textsc{full}}}
\newcommand{\manystepsto}{\stepsto^\star}

@@

The {\em result} of a machine state $\mach$ with respect to a step function
$\mathord{\stepsto} \in \MACHS \rightharpoonup \MACHS$ is a state $\mach_n$
such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step any
further.

The {\em full} step function, $\mathord{\stepstounderfull} \in \MACHS
\rightharpoonup \MACHS$, embodies the standard RISC-V ISA semantics.

@@

@@

A {\em contour}
$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}$ is a
map from machine components to ``security levels,'' where $\HIGH$ marks
locations containing secret or low-integrity data and $\LOW$ marks those
with public or high-integrity data.

Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
components agree with $\mach$ on all components marked $\LOW$ by $\contour$.
\iftext\bcp{Spell it out formally.}\fi

Formally, a {\em policy}
$\polname \in \CONTOURS \rightarrow (\MACHS \rightarrow \MACHS)$ is a
function from contours to step functions (partial functions on $\MACHS$)
that, wherever they are defined, agree with $\stepstounderfull$.  That is,
it represents an enforcement mechanism that sometimes halts programs earlier
than the bare step relation does, but that agrees with the bare step
relation when it allows evaluation to proceed.  We write
$\mach_1 \stepstounder{\pol{\contour}} \mach_2$ for
$\pol{\contour}(\mach_1) = \mach_2$.

If $\mach$ and $\mach'$ are variant machine states with respect to contour
$\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
under policy $\polname$ if either (1) one or both of $\mach$ and $\mach'$
fail to terminate when executed with $\stepstounder{\pol{\contour}}$, or (2)
both terminate and their results are variants with respect to $\contour$.

Policy $\polname$ {\em guarantees noninterference} if, for every $\contour$,
every pair $(\mach,\mach')$ of variant machine states with respect to
$\contour$ are indistinguishable under $\polname$.

@@

\bcp{I'm having some trouble here.
%
  Since the ``high part'' of the world keeps getting extended as
  tainted data gets written here and there, the only way for something
  ``bad'' to happen is for high/tainted data to get written to someplace
  that was designated as an ``output''.  We don't currently have any way of
  making such a designation.
%
  I guess we can add a ``middle'' classification level, and doing that here
  seems to make pretty good sense, but it's hard to justify adding it to the
  general definition of NI.  :-(
}

>>>>

% Notational conventions
\newcommand{\HIGH}{\textsc{h}}
\newcommand{\LOW}{\textsc{l}}
\newcommand{\OBS}{\textsc{obs}}
\newcommand{\HIDDEN}{\textsc{hidden}}
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\PCname(#1)}
\newcommand{\PCname}{\textsc{pc}}
\newcommand{\pol}{P}
\newcommand{\contour}{C}
\newcommand{\CONTOURS}{{\mathcal C}}
\newcommand{\component}{K}
\newcommand{\COMPONENTS}{{\mathcal K}}

\newcommand{\stepsto}{\Longrightarrow}
\newcommand{\stepstounder}[1]{\stackrel{\mbox{\tiny{$#1$}}}{\Longrightarrow}}
\newcommand{\stepstounderfull}{\stepstounder{\textsc{RISCV}}}
\newcommand{\manystepsto}{\stepsto^\star}
\newcommand{\obstrace}{\mathit{obstrace}}

\newcommand{\underscore}{\mbox{\_}}

@@

The {\em underlying} step function, $\mathord{\stepstounderfull} \in \MACHS
\rightharpoonup \MACHS$, embodies the standard RISC-V ISA semantics.

@@

\subsection{Policies}

A {\em contour}
$\contour \in \CONTOURS = \COMPONENTS \rightarrow \{\HIGH,\LOW\}\times
\{\OBS,\HIDDEN\}$ is a map from machine components to pairs of ``security
levels,'' where $\HIGH$ marks locations containing secret or low-integrity
data and $\LOW$ marks those with public or high-integrity data, and
``observabilities,'' where $\OBS$ marks locations whose contents are
observable from the outside (by the ``attacker'') and $\HIDDEN$ marks
internal ``hidden'' state that cannot be observed directly.

Given a machine state $\mach$ and a contour $\contour$, a {\em variant} of
$\mach$ with respect to $\contour$ is any machine state $\mach'$ whose
components agree with $\mach$ on all components marked $\LOW$ by
$\contour$---i.e., if $\contour(\component) = (\LOW$,\,\underscore), then
$\mach(\component) = \mach'(\component)$.

Formally, a {\em policy}
$\pol \in (\CONTOURS \times \MACHS) \rightarrow (\CONTOURS \times
\MACHS)$ is an enriched
step function that works on pairs of contours and machines---i.e., it is a
partial function taking a contour and a machine state to an updated contour
and machine state that, wherever is is defined, agrees with
$\stepstounderfull$.  That is,
it represents an enforcement mechanism that sometimes halts programs earlier
than the bare step relation does, but that agrees with the bare step
relation when it allows evaluation to proceed.  We write
$(\contour_1,\mach_1) \stepstounder{\pol} (\contour_2,\mach_2)$ for
$\pol(\contour_1,\mach_1) = (\contour_2,\mach_2)$.

The {\em observation trace} arising from an infinite sequence of contours
and machine states $(\contour_0,mach_0),\, (\contour_1,\mach_1),\ ldots$,
written $\obstrace((\contour_0,mach_0),\, (\contour_1,\mach_1),\ ldots)$
 is
the (finite or infinite) sequence of {\em distinct} values taken on by the
{\em observable} components of each state.  \iftext\bcp{Give an example and
  write it out formally.}\fi \bcp{Question: This could be defined in two
  ways, and I'm not sure which is better or if it really matters: the
  observation trace is a sequence of partial maps from components to their
  values, or the observation trace is a partial maps from components to
  their sequences of observable values.  This starts getting a bit
  technically tangled, but I don't see a simpler alternative right now.}

@@

\bcp{Proposal: sequences of values of the observable locations are what gets
  compared.  Fix the following:}

If $\mach$ and $\mach'$ are variant machine states with respect to contour
$\contour$, then $\mach$ and $\mach'$ are said to be {\em indistinguishable}
under policy $\pol$ if either (1) one or both of $(\contour,\mach)$ and
$(\contour,\mach')$ fail to terminate when executed with
$\stepstounder{\pol}$, or (2) both terminate and the machine-state
components of their results are variants with respect to the original
$\contour$.

Policy $\pol$ {\em guarantees noninterference} if, for every $\contour$,
every pair $(\mach,\mach')$ of variant machine states with respect to
$\contour$ are indistinguishable under $\pol$.

@@

\bcp{This needs to be written, now that we have added the hidden/observable stuff.}
